---
title: "Middle Meningeal Artery Embolization in Chronic Subdural Hematoma: A Systematic Review and Frequentist–Bayesian Meta-Analysis of Randomised Controlled Trials"
author: "Joseph Yvan Bena Nnang"
format: html
editor_options: 
  chunk_output_type: inline
---

# Packages

```{r}
if (!require("devtools")) install.packages("devtools")
if (!require("pacman")) install.packages("pacman")

if (!requireNamespace("BayesianPriorsTool", quietly = TRUE)) {
  devtools::install_github("arthur-albuquerque/BayesianPriorsTool")
}

other_packages <- c("bayesmeta", "BayesianPriorsTool", "dplyr", "tidyr", 
                    "metafor", "ggdist", "ggplot2", "patchwork", "gt", 
                    "readxl", "rstudioapi")

meta_version <- if (requireNamespace("meta", quietly = TRUE)) {
  packageVersion("meta")
} else {
  "0.0-0"
}

if (meta_version > "7.0-0") {
  remotes::install_version("meta", "7.0-0")
  pacman::p_load(char =  c("meta", other_packages))
} else if (meta_version == "7.0-0") {
  pacman::p_load(char = c("meta", other_packages))
} else {
  remotes::install_version("meta", "7.0-0")
  pacman::p_load(char = c("meta", other_packages))
}

```

```{r}
theme_set(
  theme(
    plot.title.position = 'plot',
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(size = 14),
    panel.background = element_blank(),
    panel.grid.major.x = element_line(color = "gray80", linewidth = 0.3)
  )
)

```

###Choose your output folder
```{r}
output <- selectDirectory()
```


### Configuration Switch: Define Active Outcome

```{r}

active_outcome_key <- "mrs_0_2"  # options: "mrs_0_2", "mortality", "recur"

outcome_config <- switch(active_outcome_key,
  "recur" = list(
    name = "Hematoma recurrence",
    turner_prior = "adverse events",
    direction = "less",
    enthusiastic_mean = -0.7,
    favors_left = "Favors MMAE",
    favors_right = "Favors SOC"
  ),
  "mrs_0_2" = list(
    name = "Functional Independence (mRS 0-2)",
    turner_prior = "quality of life / functioning (dichotomized)",
    direction = "greater",
    enthusiastic_mean = 0.7,
    favors_left = "Favors SOC",
    favors_right = "Favors MMAE"
  ),
  "mortality" = list(
    name = "90-Day Mortality",
    turner_prior = "all-cause mortality",
    direction = "less",
    enthusiastic_mean = -0.7,
    favors_left = "Favors MMAE",
    favors_right = "Favors SOC"
  ),
  "complications" = list(
    name = "Procedural complications",
    turner_prior = "surgical / device related success / failure",
    direction = "less",
    enthusiastic_mean = -0.7,
    favors_left = "Favors MMAE",
    favors_right = "Favors SOC"
  ),
  list(
    name = "Default Outcome",
    turner_prior = "adverse events",
    direction = "less",
    enthusiastic_mean = -0.7,
    favors_left = "Favors MMAE",
    favors_right = "Favors SOC"
  )
)

active_outcome_name <- outcome_config$name
active_turner_category <- outcome_config$turner_prior

```

### Data Loading & Effect Size Estimation

```{r}
myfile <- selectFile(caption = "choose data")
mwd <- dirname(myfile)
sheet_names <- excel_sheets(myfile)
  
ma <- lapply(sheet_names, function(x) {as.data.frame(read_excel(myfile, sheet = x))})
names(ma) <- sheet_names

data_active <- ma[[active_outcome_key]]

data_active_OR = escalc(measure="OR",
                ai=event.e, n1i=n.e, ci=event.c, n2i=n.c,
                data=data_active) |> data.frame()

data_active_OR$sei = sqrt(data_active_OR$vi)

```

### Priors Definition

```{r}
# Mu Priors
prior_mu_sceptical_mean = 0
prior_mu_sceptical_sd = 0.5

prior_mu_enthusiastic_mean = outcome_config$enthusiastic_mean
prior_mu_enthusiastic_sd = 0.5

# Tau Priors: (i) Informative (Turner et al.)
informative_tau_prior_OR =
  TurnerEtAlPrior(outcome = active_turner_category, 
                comparator1 = "non-pharmacological",
                comparator2 = "placebo / control")

# Tau Priors: (ii) Weakly Informative (Half-Cauchy(0, 0.5))
weakly_prior_OR_scale = 0.5
weakly_tau_OR_prior = function(x) {
  # Half-Cauchy PDF
  ifelse(x < 0, 0, 2 / (pi * weakly_prior_OR_scale * (1 + (x / weakly_prior_OR_scale)^2)))
}

```

### Fitting the 4 Models according to the Analysis Plan

```{r}
# (i) Primary / Sceptical: mu ~ N(0, 0.5^2), tau ~ Informative (Turner)
meta_active_primary =
  bayesmeta(y = data_active_OR$yi, sigma = data_active_OR$sei, labels = data_active_OR$author,
            mu.prior.mean = prior_mu_sceptical_mean, mu.prior.sd = prior_mu_sceptical_sd,
            tau.prior = informative_tau_prior_OR$dprior)

# (ii) Weakly Informative tau: mu ~ N(0, 0.5^2), tau ~ Half-Cauchy(0, 0.5)
meta_active_weakly =
  bayesmeta(y = data_active_OR$yi, sigma = data_active_OR$sei, labels = data_active_OR$author,
            mu.prior.mean = prior_mu_sceptical_mean, mu.prior.sd = prior_mu_sceptical_sd,
            tau.prior = weakly_tau_OR_prior)

# (iii) Enthusiastic prior: mu ~ N(-0.7, 0.5^2), tau ~ Informative (Turner)
meta_active_enthusiastic =
  bayesmeta(y = data_active_OR$yi, sigma = data_active_OR$sei, labels = data_active_OR$author,
            mu.prior.mean = prior_mu_enthusiastic_mean, mu.prior.sd = prior_mu_enthusiastic_sd,
            tau.prior = informative_tau_prior_OR$dprior)

```

### Primary Analysis Visualizations

```{r}
png(file = paste0(output,"/","Overall effect ", active_outcome_name, ".png"),
    width = 3000, height = 2500, res = 300)
plot(meta_active_primary, which = 3, main = "Overall Effect (Primary/Sceptical)", prior = TRUE)                             
dev.off() 

png(file = paste0(output,"/","Between study heterogeneity ", active_outcome_name, ".png"),
    width = 3000, height = 2500, res = 300)
plot(meta_active_primary, which = 4, main = "Between-study Heterogeneity", prior = TRUE)                                   
dev.off() 

```

### ------------------------------------------------------------------
### Primary Analysis: Native Bayesian Forest Plot
### ------------------------------------------------------------------

```{r}

# 1. Extract posterior summary statistics from your primary Bayesian model
bayes_mu_mean  <- meta_active_primary$summary["mean", "mu"]
bayes_mu_lower <- meta_active_primary$summary["95% lower", "mu"]
bayes_mu_upper <- meta_active_primary$summary["95% upper", "mu"]

# Extract median between-study heterogeneity (tau) and convert to character label
bayes_tau_median <- round(meta_active_primary$summary["median", "tau"], 3)
tau_label_text   <- paste0("Heterogeneity (Tau) = ", bayes_tau_median)

# 2. Inject these values into your frequentist object to overwrite its summary line
meta_forest_primary_revman <- metaadd(
  forest_data_active, 
  type  = "random", 
  TE    = bayes_mu_mean, 
  lower = bayes_mu_lower, 
  upper = bayes_mu_upper
)

# 3. Open PNG device
png(file = paste0(output, "/", "Bayesian Forest Plot Primary ", active_outcome_name, ".png"),
    width = 4500, height = 2500, res = 300)

# 4. Generate the structured RevMan5 plot using the modified object
meta::forest(
  meta_forest_primary_revman,
  layout = "RevMan5",
  sortvar = TE,           # Sorts cleanly by raw study effect size
  text.random = "Bayesian Posterior",    # Relabels the summary diamond line
  text.addline1 = tau_label_text,        # Dynamic, error-free heterogeneity text
  smlab = "Odds Ratio",
  
  # Column Configurations
  leftcols = c("studlab", "event.e", "n.e", "event.c", "n.c", "effect.ci"),
  leftlabs = c("Study", "Events", "Total", "Events", "Total", "OR (95% CrI)"),
  label.left = outcome_config$favors_left, 
  label.right = outcome_config$favors_right,
  pooled.events = TRUE,
  lab.e = "MMAE", 
  lab.c = "SOC",
  colgap = "5mm",
  common = FALSE,                        # Strips the fixed-effect diamond out
  random = TRUE                          # Keeps your forced Bayesian diamond visible
)

dev.off()
```


### Sensitivity Analysis: Forest Plots combining all Priors

```{r}
# Base frequentist data format for metaadd
forest_data_active = metabin(
    data = data_active_OR, studlab = author,
    event.e = event.e, n.e = n.e, event.c = event.c, n.c = n.c, 
    sm = "OR", common = FALSE, random = FALSE
)

# Extract Posteriors
extract_posterior <- function(meta_obj) {
  list(
    mu_mean = meta_obj$summary["mean", "mu"],
    mu_upper = meta_obj$summary["95% upper", "mu"],
    mu_lower = meta_obj$summary["95% lower", "mu"],
    theta_mean = meta_obj$summary["mean", "theta"],
    theta_upper = meta_obj$summary["95% upper", "theta"],
    theta_lower = meta_obj$summary["95% lower", "theta"]
  )
}

post_primary <- extract_posterior(meta_active_primary)
post_weakly <- extract_posterior(meta_active_weakly)
post_enthusiastic <- extract_posterior(meta_active_enthusiastic)

# Build metaadd objects for Overall Effect (mu)
meta_forest_primary = metaadd(forest_data_active, type = "random", 
                              TE = post_primary$mu_mean, upper = post_primary$mu_upper, lower = post_primary$mu_lower)  
meta_forest_weakly = metaadd(forest_data_active, type = "random", 
                             TE = post_weakly$mu_mean, upper = post_weakly$mu_upper, lower = post_weakly$mu_lower)  
meta_forest_enthusiastic = metaadd(forest_data_active, type = "random", 
                                   TE = post_enthusiastic$mu_mean, upper = post_enthusiastic$mu_upper, lower = post_enthusiastic$mu_lower)  


# Combine for Plot (Dynamically reading the enthusiastic mean)
analyses_labels = c(
  "(i) Sceptical / Primary prior [mu ~ N(0, 0.5²), tau ~ Turner]",
  "(ii) Weakly Informative prior [mu ~ N(0, 0.5²), tau ~ Half-Cauchy(0, 0.5)]",
  paste0("(iii) Enthusiastic prior [mu ~ N(", prior_mu_enthusiastic_mean, ", 0.5²), tau ~ Turner]")
)

all_active_analyses <- metabind(meta_forest_primary,
                                meta_forest_weakly,
                                meta_forest_enthusiastic,
                                name = analyses_labels)

png(file = paste0(output,"/","Forest plot sensitivity ", active_outcome_name, ".png"),
    width = 4500, height = 3000, res = 300)

meta::forest(all_active_analyses, 
             smlab = "Overall Effect", 
             leftcols ="studlab",
             pooled.events = TRUE,
             lab.e = "MMAE", lab.c = "SOC",
             label.left = outcome_config$favors_left, 
             label.right = outcome_config$favors_right,
             colgap = "5mm")
dev.off() 

```

### Sensitivity Analysis: Prediction Intervals

```{r}
# Build metaadd objects for Prediction Intervals (theta)
meta_pred_primary = metaadd(forest_data_active, type = "prediction", 
                            TE = post_primary$theta_mean, upper = post_primary$theta_upper, lower = post_primary$theta_lower)  
meta_pred_weakly = metaadd(forest_data_active, type = "prediction", 
                           TE = post_weakly$theta_mean, upper = post_weakly$theta_upper, lower = post_weakly$theta_lower)  
meta_pred_enthusiastic = metaadd(forest_data_active, type = "prediction", 
                                 TE = post_enthusiastic$theta_mean, upper = post_enthusiastic$theta_upper, lower = post_enthusiastic$theta_lower)  

all_pred_analyses <- metabind(meta_pred_primary,
                              meta_pred_weakly,
                              meta_pred_enthusiastic,
                              name = analyses_labels)

png(file = paste0(output,"/","Prediction intervals sensitivity ", active_outcome_name, ".png"),
    width = 4500, height = 3000, res = 300)

meta::forest(all_pred_analyses, 
             smlab = "Prediction Intervals", 
             leftcols ="studlab",
             label.left = outcome_config$favors_left, 
             label.right = outcome_config$favors_right
             )
dev.off() 

```



### ------------------------------------------------------------------
### Posterior Probabilities of Benefit 
### ------------------------------------------------------------------

```{r}
# Determine whether we are calculating OR < 1 or OR > 1 based on the outcome type
if (outcome_config$direction == "less") {
  
  # Probability of risk reduction (OR < 1)
  prob_primary      = meta_active_primary$pposterior(mu = log(1))
  prob_weakly       = meta_active_weakly$pposterior(mu = log(1))
  prob_enthusiastic = meta_active_enthusiastic$pposterior(mu = log(1))
  display_label     = "OR < 1"
  
} else {
  
  # Probability of clinical improvement (OR > 1) -> 1 minus the cumulative probability below 0
  prob_primary      = 1 - meta_active_primary$pposterior(mu = log(1))
  prob_weakly       = 1 - meta_active_weakly$pposterior(mu = log(1))
  prob_enthusiastic = 1 - meta_active_enthusiastic$pposterior(mu = log(1))
  display_label     = "OR > 1"
  
}

cat("--- Posterior Probabilities of Clinical Benefit (", display_label, ") ---\n", sep = "")
cat("Outcome: ", active_outcome_name, "\n")
cat("(i) Primary/Sceptical Model: ", round(prob_primary * 100, 2), "%\n")
cat("(ii) Weakly Informative Model:   ", round(prob_weakly * 100, 2), "%\n")
cat("(iii) Enthusiastic Model:         ", round(prob_enthusiastic * 100, 2), "%\n\n")

```
```{r}
# Determine whether we are calculating OR < 1 or OR > 1 based on the outcome type
if (outcome_config$direction == "less") {
  
  # Probability of risk reduction (OR < 1)
  prob_primary      = meta_active_primary$pposterior(mu = log(0.5))
  prob_weakly       = meta_active_weakly$pposterior(mu = log(0.5))
  prob_enthusiastic = meta_active_enthusiastic$pposterior(mu = log(0.5))
  display_label     = "OR < 0.5"
  
} else {
  
  # Probability of clinical improvement (OR > 1) -> 1 minus the cumulative probability below 0
  prob_primary      = 1 - meta_active_primary$pposterior(mu = log(0.5))
  prob_weakly       = 1 - meta_active_weakly$pposterior(mu = log(0.5))
  prob_enthusiastic = 1 - meta_active_enthusiastic$pposterior(mu = log(0.5))
  display_label     = "OR > 0.5"
  
}

cat("--- Posterior Probabilities of Clinical Benefit (", display_label, ") ---\n", sep = "")
cat("Outcome: ", active_outcome_name, "\n")
cat("(i) Primary/Sceptical Model: ", round(prob_primary * 100, 2), "%\n")
cat("(ii) Weakly Informative Model:   ", round(prob_weakly * 100, 2), "%\n")
cat("(iii) Enthusiastic Model:         ", round(prob_enthusiastic * 100, 2), "%\n\n")

```



```{r}
# Determine whether we are calculating OR < 1 or OR > 1 based on the outcome type
if (outcome_config$direction == "less") {
  
  # Probability of risk reduction (OR < 1)
  prob_primary      = meta_active_primary$pposterior(mu = log(0.75))
  prob_weakly       = meta_active_weakly$pposterior(mu = log(0.75))
  prob_enthusiastic = meta_active_enthusiastic$pposterior(mu = log(0.75))
  display_label     = "OR < 0.75"
  
} else {
  
  # Probability of clinical improvement (OR > 1) -> 1 minus the cumulative probability below 0
  prob_primary      = 1 - meta_active_primary$pposterior(mu = log(0.75))
  prob_weakly       = 1 - meta_active_weakly$pposterior(mu = log(0.75))
  prob_enthusiastic = 1 - meta_active_enthusiastic$pposterior(mu = log(0.75))
  display_label     = "OR > 0.75"
  
}

cat("--- Posterior Probabilities of Clinical Benefit (", display_label, ") ---\n", sep = "")
cat("Outcome: ", active_outcome_name, "\n")
cat("(i) Primary/Sceptical Model: ", round(prob_primary * 100, 2), "%\n")
cat("(ii) Weakly Informative Model:   ", round(prob_weakly * 100, 2), "%\n")
cat("(iii) Enthusiastic Model:         ", round(prob_enthusiastic * 100, 2), "%\n\n")

```



### ------------------------------------------------------------------
### Posterior Distribution Visualizations (Density Plots)
### ------------------------------------------------------------------

```{r}
# Extracting samples from the posterior distribution for visualization
set.seed(123)
n_samples = 10000

samples_primary = meta_active_primary$rposterior(n = n_samples)[,"mu"]
samples_weakly = meta_active_weakly$rposterior(n = n_samples)[,"mu"]
samples_enthusiastic = meta_active_enthusiastic$rposterior(n = n_samples)[,"mu"]

# Compiling into a single dataframe for ggplot
posterior_data = bind_rows(
  data.frame(Model = "(i) Sceptical / Primary", Value = samples_primary),
  data.frame(Model = "(ii) Weakly Informative", Value = samples_weakly),
  data.frame(Model = "(iii) Enthusiastic", Value = samples_enthusiastic)
)

# Reordering factor levels so they plot in a logical sequence
posterior_data$Model = factor(posterior_data$Model, 
                              levels = c("(iii) Enthusiastic", 
                                         "(ii) Weakly Informative", 
                                         "(i) Sceptical / Primary"))

# Generating the Plot using ggdist
plot_posteriors = ggplot(posterior_data, aes(x = Value, y = Model, fill = Model)) +
  stat_halfeye(alpha = 0.7, .width = c(0.95), point_interval = "median_qi") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray30", linewidth = 0.8) +
  labs(
    title = paste("Posterior Distributions of Overall Effect -", active_outcome_name),
    subtitle = "Comparing Sceptical, Weakly Informative, and Enthusiastic Priors",
    x = "Log Odds Ratio (log-OR)",
    y = NULL
  ) +
  scale_fill_manual(values = c("(iii) Enthusiastic" = "#4DB6AC", 
                               "(ii) Weakly Informative" = "#7986CB", 
                               "(i) Sceptical / Primary" = "#E57373")) +
  theme(legend.position = "none")

# Save the plot
png(file = paste0(output,"/","Posterior Distributions ", active_outcome_name, ".png"),
    width = 3500, height = 2500, res = 300)
print(plot_posteriors)
dev.off()
```


### Savage-Dickey Density Ratio Computation (Primary/Sceptical Model)

```{r}
null_point <- 0 

# Primary / Sceptical Model
prior_dens_sceptical <- dnorm(null_point, mean = prior_mu_sceptical_mean, sd = prior_mu_sceptical_sd)
post_dens_sceptical  <- meta_active_primary$dposterior(mu = null_point)
BF01_sceptical       <- post_dens_sceptical / prior_dens_sceptical
BF10_sceptical       <- prior_dens_sceptical / post_dens_sceptical

# Enthusiastic Model
prior_dens_enthusiastic <- dnorm(null_point, mean = prior_mu_enthusiastic_mean, sd = prior_mu_enthusiastic_sd)
post_dens_enthusiastic  <- meta_active_enthusiastic$dposterior(mu = null_point)
BF01_enthusiastic       <- post_dens_enthusiastic / prior_dens_enthusiastic
BF10_enthusiastic       <- prior_dens_enthusiastic / post_dens_enthusiastic

cat("--- Savage-Dickey Bayes Factor Results ---\n")
cat("Outcome: ", active_outcome_name, "\n\n")
cat("[Primary/Sceptical Prior Model]\n")
cat("Bayes Factor BF10 (H1 vs H0): ", round(BF10_sceptical, 4), "\n")
cat("Bayes Factor BF01 (H0 vs H1): ", round(BF01_sceptical, 4), "\n\n")
cat("[Enthusiastic Prior Model]\n")
cat("Bayes Factor BF10 (H1 vs H0): ", round(BF10_enthusiastic, 4), "\n")
cat("Bayes Factor BF01 (H0 vs H1): ", round(BF01_enthusiastic, 4), "\n")

```

# Conceptual structure

```{r}
# Save the plot
png(file = paste0(output,"/","Conceptual structure ", active_outcome_name, ".png"),
    width = 3500, height = 2500, res = 300)

ggplot() +
  stat_function(fun = dnorm, args = list(mean=0, sd=0.5), aes(color="Prior")) +
  geom_density(data=data.frame(x=samples_primary), aes(x, color="Posterior")) +
  geom_vline(xintercept=0) # null point
  
  dev.off()
```

# 1. Compile the calculated probabilities into a structured dataframe

```{r}
prob_summary_data <- data.frame(
  Model = c("(i) Sceptical / Primary", 
            "(ii) Weakly Informative", 
            "(iii) Enthusiastic"),
  Probability = c(prob_primary, prob_weakly, prob_enthusiastic) * 100
)

# 2. Maintain strict factor level ordering to match the density plots
prob_summary_data$Model <- factor(
  prob_summary_data$Model, 
  levels = c("(iii) Enthusiastic", 
             "(ii) Weakly Informative", 
             "(i) Sceptical / Primary")
)

# 3. Construct the Horizontal Bar Chart
plot_probabilities <- ggplot(prob_summary_data, aes(x = Probability, y = Model, fill = Model)) +
  geom_col(width = 0.6, alpha = 0.85, color = "white", linewidth = 0.5) +
  # Add inline text percentage labels next to the bars
  geom_text(aes(label = paste0(round(Probability, 2), "%")), 
            hjust = -0.15, size = 4.5, fontface = "bold", color = "gray20") +
  # Enforce a 0-100% axis layout with extra padding for labels
  scale_x_continuous(limits = c(0, 115), breaks = seq(0, 100, 20), expand = c(0, 0)) +
  labs(
    title = paste("Posterior Probability of Clinical Benefit:", active_outcome_name),
    subtitle = paste0("Threshold Criterion: Probability that true treatment effect satisfies ", display_label),
    x = paste0("Probability (%, ", display_label, ")"),
    y = NULL
  ) +
  scale_fill_manual(values = c("(iii) Enthusiastic" = "#4DB6AC", 
                               "(ii) Weakly Informative" = "#7986CB", 
                               "(i) Sceptical / Primary" = "#E57373")) +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_line(color = "gray90", linewidth = 0.4),
    axis.text.y = element_text(face = "bold", size = 11)
  )

# 4. Save high-resolution graphic asset for manuscript submission
png(file = paste0(output, "/", "Posterior Probability Summary ", active_outcome_name, ".png"),
    width = 3500, height = 2000, res = 300)
print(plot_probabilities)
dev.off()

```
